---
title             : "Music Emotion Recognition: Meta-analysis"
shorttitle        : "XXXXXXXXX"

author: 
  - name          : "TE"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Palace Green, Department of Music, Durham University, DH1 3RL Durham, United Kingdom"
    email         : "tuomas.eerola@durham.ac.uk"
  - name          : "CA"
    affiliation   : "2"
    address       : "XXXXXX"
    email         : "XXX@XXXX"

affiliation:
  - id            : "1"
    institution   : "Department of Music, Durham University"
  - id            : "2"
    institution   : "XXXX, MacMaster University"

note: ""

author_note: |
  XXX.

abstract: |
 blah blah blah.

keywords          : "music, emotion, recognition, meta-analysis"
wordcount         : "XXXX"

bibliography      : ["references.bib"]

figsintext        : yes
figurelist        : no
tablelist         : no
footnotelist      : no
lineno            : no
mask              : no
csl: apa7.csl

class             : "man"
output            : papaja::apa6_pdf
---

```{r global options, include = FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
```

```{r include = FALSE, echo=FALSE}
#source('load_libraries.R')
## set options ##
```


# Introduction

blah blah [@anderson2022ex]


## Aims

blah blah

# Methods

blah blah

We preregistered our meta-analysis plan (OSF URL here and available also at [https://tuomaseerola.github.io/metaMER/preregistration/preregistration.html](https://tuomaseerola.github.io/metaMER/preregistration/preregistration.html)).

# Results

blah blah

# Conclusion and Discussion

blah blah

### Funding statement

The authors... 

### Competing interests statement

There were no competing interests.

### Open practices statement

XXXX. 

### Acknowledgements

xxxxx.

# References

